SIG Proceedings Paper in LaTeX Format

### 2019-09-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/background.tex
bib modifiactions plus background and introduction review

\fi
This section gives in its first subsection a broad overview over the functioning of the Skip-Gram Model, and its second section describes the Technique called Negative Sampling introduced by Mikolov et al. \citep{mikolov2} used to make the Skip-Gram model more efficient
\subsection{The Skip-Gram Model}
The SGM is a very simple model used to learn WE. The idea is to train a neural network, on a fake task and then use some of the weights of the network as embeddings. To understand this fake task center and context words must be defined. The center word is any given word in a sentence, from which we want to learn the WE. The context words of this specific center word are the words left and right in a given window $m$ in the sentence. See Figure \ref{fig:window_ex} as an example, where the context words are highlighted. An important aspect to notice, is that instead of having a fixed window size $m$, each word will randomly have a window size between $1$ and $m$. The idea behind having different window sizes is that words that are further away in the sentence, have less semantic correlation to the center word.

@@

\subsection{Negative Sampling}
A second alternative solution to the use of a classical softmax is negative sampling. Negative Sampling is a simplification of Noise Contrastive Estimation (NCE) which was introduced by Gutmann and Hyv{\"a}rinen \citep{nce-original}, and first applied to NLP by Mnih and Teh \citep{mnih}. This subsection will shortly describe NCE and then how Mikolov et al. \citep{mikolov2} simplified it to create the technique called negative sampling. \\ The idea behind NCE is to distinguish data from noise. It does so by reducing the problem to a logistic regression task and does it by maximizing the log probability. The SGM is only interested in good word representation, hence the probability of the word is not meaningful as long as the quality of the word representations remains high. Mikolov et al. \citep{mikolov2} simplified NCE and called it Negative Sampling, which will be explained next.\\
The main goal of negative sampling is to only update the output nodes of certain words, 5-20 in practice. This will obviously save an enormous amount of computation time. The idea is that given a pair $(c,w) \in D$, wh$w$ ere $c$ is a word in the context window of we select $K$ random words $k_i$ from the corpus data $D$. We assume those words do not appear in the context of $w$. By doing so we will only have to update $k+1$ nodes. Mikolov et al. \citep{mikolov2} introduced the following objective function $J$, that should be maximized:

@@

Where $v_c$ and $\tilde{v_w }$, can be interpreted as before in Equation \ref{eq:softmax} and $\sigma(x) = \frac{1}{1+e^{-x}}$. To  understand why this loss function is working so well, one needs to assume that two vectors are similar if their dot product is high. Therefore, we maximize the dot product of similar words $(w,c)$ ($c$ appears in the context of $w$) and minimize it for dissimilar words $(w,k_i)$ (those were selected randomly, we therefore assume that their are not similar).
We see that to compute our objective function we will only have to compute the sum over the number of negative samples $K$, which is very small in practice (2-20). To put things in perspective lets imagine our data set consists of 100000 words, we set $K=2$. Assume each output neuron has a weight vector of dimension $d = 300$. In consequence, when updating our weights we would only update $0.2*10^{-2}$ of the 300 million weights in the output layer.

@@

\begin{equation} \label{eq:unigram}
P(w)=\frac{f(w)^{\frac{3}{4}}}{\sum_{t=0}^{T} f(w_t)^{\frac{3}{4}}}
\end{equation}
where $f(w)$ is the frequency of $w_t$. The value of $\frac{3}{4}$ is set empirically. Raising the unigram distribution to the power of $\frac{3}{4}$ makes it less likely for a word to be drawn if it often appears in the dataset in comparison to the basic unigram distribution.
\iffalse See figure \ref{fig:frequency_ex} for an example.\\
\begin{figure}[ht]
\centering
\includegraphics[scale=0.30]{images/frequency_ex}
\caption{Probabililty of a word, of the text8 dataset (sampled), to be chosen at random according to its frequency and the power to which the unigram distribution is raised}
\label{fig:frequency_ex}
\end{figure}
\fi
With all of the above knowledge it's understandable that the SGNS will  will outperform the classical softmax in computation time. Now the question arises if the accuracy is maintained but according to Mikolov et al. \citep{mikolov2} the negative sampling method "is an extremely simple training method that learns accurate representations". As a matter of fact, Mikolov et al. \citep{mikolov2} reported a 6\% accuracy improvement in comparison to a Hierarchical Softmax model. Therefore, it is a good solution to the problem raised by the classical softmax.
We now have enough background knowledge about the SGM to look at how it can be optimized. The next section introduces earlier approaches to optimize the SGM.

>>>>

\fi
This section gives in its first subsection a broad overview over the functioning of the Skip-Gram Model, and its second section describes the technique called Negative Sampling introduced by Mikolov et al. \citep{mikolov2} used to make the Skip-Gram model more efficient.
\subsection{The Skip-Gram Model}
The SGM is a very simple model used to learn WE. The idea is to train a neural network, on a fake task and then use some of the weights of the network as embeddings. To understand this fake task center and context words must be defined. The center word is any given word in a sentence, from which we want to learn the WE. The context words of this specific center word are the words left and right in a given window $m$ in the sentence. See Figure \ref{fig:window_ex} as an example, where the context words are highlighted. An important aspect to notice, is that instead of having a fixed window size $m$, each word will randomly have a window size between $1$ and $m$. The idea behind having different window sizes is that words that are further away in the sentence, have less semantic correlation to the center word.

@@

\subsection{Negative Sampling}
A second alternative solution to the use of a classical softmax is negative sampling. Negative Sampling is a simplification of Noise Contrastive Estimation (NCE) which was introduced by Gutmann and Hyv{\"a}rinen \citep{nce-original}, and first applied to NLP by Mnih and Teh \citep{mnih}. This subsection will shortly describe NCE and then how Mikolov et al. \citep{mikolov2} simplified it to create the technique called negative sampling. \\ The idea behind NCE is to distinguish data from noise. It does so by reducing the problem to a logistic regression task and does it by maximizing the log probability. The SGM is only interested in good word representation, hence the probability of the word is not meaningful as long as the quality of the word representations remains high. Mikolov et al. \citep{mikolov2} simplified NCE and called it Negative Sampling, which will be explained next.\\
The main goal of negative sampling is to only update the output nodes of certain words, 5-20 in practice. This will obviously save an enormous amount of computation time. The idea is that given a pair $(c,w) \in D$, where $w$  $c$ is a word in the context window of $c$ we select $K$ random words $k_i$ from the corpus data $D$. We assume those words do not appear in the context of $w$. By doing so we will only have to update $k+1$ nodes. Mikolov et al. \citep{mikolov2} introduced the following objective function $J$, that should be maximized:

@@

Where $v_c$ and $\tilde{v_w }$, can be interpreted as before in Equation \ref{eq:softmax} and $\sigma(x) = \frac{1}{1+e^{-x}}$. To  understand why this loss function is working so well, one needs to assume that two vectors are similar if their dot product is high. Therefore, we maximize the dot product of similar words $(w,c)$ ($c$ appears in the context of $w$) and minimize it for dissimilar words $(w,k_i)$ (those were selected randomly, we therefore assume that their are not similar).
We see that to compute our objective function we will only have to compute the sum over the number of negative samples $K$, which is very small in practice (2-20). To put things in perspective lets imagine our data set consists of 100000 words, we set $K=2$. Assume each output neuron has a weight vector of dimension $d = 300$. In consequence, when updating our weights we would only update $0.9*10^{3}$ of the 300 million weights in the output layer.

@@

\begin{equation} \label{eq:unigram}
P(w)=\frac{f(w)^{\frac{3}{4}}}{\sum_{t=0}^{T} f(w_t)^{\frac{3}{4}}}
\end{equation}
where $f(w)$ is the frequency of $w_t$. The value of $\frac{3}{4}$ is set empirically. Raising the unigram distribution to the power of $\frac{3}{4}$ makes it less likely for a word to be drawn if it often appears in the dataset in comparison to the basic unigram distribution.
\iffalse See figure \ref{fig:frequency_ex} for an example.\\
\begin{figure}[ht]
\centering
\includegraphics[scale=0.30]{images/frequency_ex}
\caption{Probabililty of a word, of the text8 dataset (sampled), to be chosen at random according to its frequency and the power to which the unigram distribution is raised}
\label{fig:frequency_ex}
\end{figure}
\fi
With all of the above knowledge it's understandable that the SGNS will  outperform the classical softmax in computation time. Now the question arises if the accuracy is maintained but according to Mikolov et al. \citep{mikolov2} the negative sampling method "is an extremely simple training method that learns accurate representations". As a matter of fact, Mikolov et al. \citep{mikolov2} reported a 6\% accuracy improvement in comparison to a Hierarchical Softmax model. Therefore, it is a good solution to the problem raised by the classical softmax.
We now have enough background knowledge about the SGM to look at how it can be optimized. The next section introduces earlier approaches to optimize the SGM.

### 2019-09-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/results.tex
bib modifiactions plus background and introduction review

\subsection{Simple Batch Version}
\subsubsection{SGD}
The first challenge for each optimizer was to find a correct learning rate. As SGD is the optimizer used in Gensim \citep{gensim} we first tried the same learning rate as the default value in Gensim \citep{gensim}, i.e 0.01,  and then performed a random search to find a better one. As expected a bell curve shape resulted, a learning rate that is too high leads to diversion and a learning rate that is too low leads to a training time that is too slow. The best value that we found for the learning rate is $0.0075$. With this setting SGD converged in 11 epochs. The second experiment was to add input shuffling.
As seen in Figure \ref{fig:results_sgd}, for almost every learning rate the convergence time decreased. Our model, with the best setting, know converges in only 7 epochs. Another interesting fact to point out from Figure \ref{fig:results_sgd} is that with input shuffling we achieved better results with higher learning rates. As for learning rates of $0.01$ and $0.025$ we did converge in 11 epochs with input shuffling but did not converge in 20 epochs without it.

\begin{figure}[h]
\centering
\includegraphics[scale=0.3]{images/results_sgd_shuffle}
\caption{Training time Stochastic Gradient Descent with input Shuffling}
\label{fig:results_sgd}
\end{figure}
\iffalse 
\subsubsection{Momentum and Nesterov}
Momentum and  Nesterov Accelerated Gradient (NAG) \citep{nag} both have an additional hyperparameter $\gamma$, that defines the percentage of the previous gradient that will be added to the current gradients. We set $\gamma = 0.9$ as this is a typical value and did not alter it during our experiments. Momentum and NAG alone respectively only slightly decrease or increase the convergence time. Momentum optimally converges in 9 epochs and NAG in 13. If we combine these optimizers with input shuffling, interestingly the same phenomena as with plain SGD appear. The convergence time gets better, 8 epochs for Momentum and 3 epochs for NAG. The phenomena that a higher learning rat yields better results also happens with both of the optimizers. As Momentum does not converge in 20 epochs with a learning rate of 0.002 but does in 8 with input shuffling.
\begin{figure}[h]
\centering

@@

\subsubsection{Adagrad}\label{ssec:results_adagrad}
Adagrad \citep{adagrad} is a very interesting optimizing for creating word embeddings as it decreases the learning rate for very frequent occurring features, and vice versa. Words (features in our case) that appear very frequently (low learning rate) often do not have a semantic gain  that is as important as words that appear less frequently (high learning rate) to their context words. So, in theory, Adagrad is particularly well suited for our task, as for example Pennington et al. used Adagrad in the training of Glove \citep{glove}, another system which word embeddings. Our results confirmed empirically that Adagrad can work better for creating word embeddings. The  model converged in 4 epochs. When combined with shuffling Adagrad only took 3 epochs to converge. These results confirms two hypothesis: first  the tendency of the Skip-Gram Model to converge faster with input shuffling and second  the big impact of having different learning rate for each feature.
Another interesting fact to notice is that a higher learning rate combined with input shuffling did not yield better results than without shuffling, in comparison to the earlier described optimzers. Adagrad performed best with a learning rate of $0.1$, as shown in Figure \ref{fig:results_adagrad_shuffle}.
\begin{figure}[h]
\centering
\includegraphics[scale=0.3]{images/results_adagrad_shuffle}
\caption{Training time Adagrad with input Shuffling}
\label{fig:results_adagrad_shuffle}
\end{figure}
\subsubsection{Adadelta}
In theory Adadelta \citep{adadelta} should outperform Adagrad as it's an extension of the former. Adadelta doesn't  have any learning rate to tune, therefore we only did 2 experiments, with and without input shuffling. 
We are aware of the fact that there are additional hyper parameters to Adadelta. We decide not to tune them for two reasons: first for simplicity reasons and second because their effect is not as high as the learning rate. The hyper parameter, $\rho$ that defines the percentage taken when calculating the exponentially decaying average of past gradients was set to $\rho = 0.9$. Adadelta did not manage to achieve a word similarity of 0.66, in comparison to all the others optimzers. It only converged to a similarity of 0.59. It did this in 20 epochs without input shuffling and in 3 with input shuffling, as can be seen in Table \ref{table:results_adadelta}

@@

\subsubsection{Adam}
Adam is the most advanced of all the optimizers used in our experiments and did yield the best results as seen in Figure \ref{fig:results_adam_shuffle}. Adam converged in 3 epochs without shuffling and a learning rate of $0.05$ and 2 with shuffling and a learning rate of $0.001$, as shown in Figure \ref{fig:results_adam_shuffle}. Those are the best results achieved in any of our experiment. Exacctly as Adagrad, Adamd did not react to input shuffling the same way as SGD did.
\begin{figure}[h]
    \centering
            \includegraphics[scale=0.3]{images/results_adam_shuffle} 
    \caption{Training time Adam with input Shuffling}
    \label{fig:results_adam_shuffle}
\end{figure}

\subsection{Advanced Batches Version}
The advanced batched version is largely dependent on the definition of outlier. As shown in Table if the value for outliers is too small then the model convergence into score that is too small. If the threshold is higher training time is increased. We found that 99\% is a good reference value. With this setting our model converged in 1 epoch for the text8 dataset, which is our best result, and 7 for the enwik9 corpus. A recap of all our results with the text8 dataset can be found in Table.

\begin{table}[tb]
    \caption{Results on the text8 dataset}
    \begin{tabular}{l r r }
        \toprule
Model    & Convergence Time & Word Similarity \\ \hline
SGD & {11}              & 0.65            \\ 
SGD w/shuffling & {7}              & 0.66            \\ 
Adam & {3}              & 0.66            \\ 
Adam w/ shuffling & {2}              & 0.66      \\ 
Gensim   & 4          & 0.66            \\ \hline
        %
\textbf{Adam w/o outliers }& \textbf{1} & 0.66 \\  \hline
\bottomrule
\end{tabular}
\end{table}

>>>>

\subsubsection{SGD}
The first challenge for each optimizer was to find a correct learning rate. As SGD is the optimizer used in Gensim \citep{gensim} we first tried the same learning rate as the default value in Gensim \citep{gensim}, i.e 0.01,  and then performed a random search to find a better one. As expected a bell curve shape resulted, a learning rate that is too high leads to diversion and a learning rate that is too low leads to a training time that is too slow. The best value that we found for the learning rate is $0.0075$. With this setting SGD converged in 11 epochs. The second experiment was to add input shuffling.
For almost every learning rate the convergence time decreased. Our model, with the best setting, know converges in only 7 epochs. Another interesting fact to point out is that with input shuffling we achieved better results with higher learning rates. As for learning rates of $0.01$ and $0.025$ we did converge in 11 epochs with input shuffling but did not converge in 20 epochs without it.
\iffalse
\begin{figure}[h]
\centering
\includegraphics[scale=0.3]{images/results_sgd_shuffle}
\caption{Training time Stochastic Gradient Descent with input Shuffling}
\label{fig:results_sgd}
\end{figure}
\fi
\iffalse 
\subsubsection{Momentum and Nesterov}
Momentum and  Nesterov Accelerated Gradient (NAG) \citep{nag} both have an additional hyperparameter $\gamma$, that defines the percentage of the previous gradient that will be added to the current gradients. We set $\gamma = 0.9$ as this is a typical value and did not alter it during our experiments. Momentum and NAG alone respectively only slightly decrease or increase the convergence time. Momentum optimally converges in 9 epochs and NAG in 13. If we combine these optimizers with input shuffling, interestingly the same phenomena as with plain SGD appear. The convergence time gets better, 8 epochs for Momentum and 3 epochs for NAG. The phenomena that a higher learning rat yields better results also happens with both of the optimizers. As Momentum does not converge in 20 epochs with a learning rate of 0.002 but does in 8 with input shuffling.
\begin{figure}[h]
\centering

@@

\subsubsection{Adadelta}
In theory Adadelta \citep{adadelta} should outperform Adagrad as it's an extension of the former. Adadelta doesn't  have any learning rate to tune, therefore we only did 2 experiments, with and without input shuffling. 
We are aware of the fact that there are additional hyper parameters to Adadelta. We decide not to tune them for two reasons: first for simplicity reasons and second because their effect is not as high as the learning rate. The hyper parameter, $\rho$ that defines the percentage taken when calculating the exponentially decaying average of past gradients was set to $\rho = 0.9$. Adadelta did not manage to achieve a word similarity of 0.66, in comparison to all the others optimzers. It only converged to a similarity of 0.59. It did this in 20 epochs without input shuffling and in 3 with input shuffling, as can be seen in Table \ref{table:results_adadelta}

@@

\subsubsection{Adagrad}\label{ssec:results_adagrad}
Adagrad \citep{adagrad} is a very interesting optimizer for creating word embeddings as it decreases the learning rate for very frequent occurring features, and vice versa. Words (features in our case) that appear very frequently (low learning rate) often do not have a semantic gain  that is as important as words that appear less frequently (high learning rate) to their context words. So, in theory, Adagrad is particularly well suited for our task, as for example Pennington et al. used Adagrad in the training of Glove \citep{glove}, another system which word embeddings. Our results confirmed empirically that Adagrad can work better for creating word embeddings. The  model converged in 4 epochs. When combined with shuffling Adagrad only took 3 epochs to converge, as shown in Table \ref{table:results}. These results confirms two hypothesis: first  the tendency of the Skip-Gram Model to converge faster with input shuffling and second  the big impact of having different learning rate for each feature.
Another interesting fact to notice is that a higher learning rate combined with input shuffling did not yield better results than without shuffling, in comparison to the earlier described optimzers. Adagrad performed best with a learning rate of $0.1$.
\iffalse
\begin{figure}[h]
\centering
\includegraphics[scale=0.3]{images/results_adagrad_shuffle}
\caption{Training time Adagrad with input Shuffling}
\label{fig:results_adagrad_shuffle}
\end{figure}
\fi
\subsubsection{Adam}
Adam is the most advanced of all the optimizers used in our experiments and did yield the best results as seen in F. Adam converged in 3 epochs without shuffling and a learning rate of $0.05$ and 2 with shuffling and a learning rate of $0.001$, as shown in Table \ref{table:results}. Those are the best results achieved in any of our experiment. Exacctly as Adagrad, Adamd did not react to input shuffling the same way as SGD did.
    \begin{table}[]
\begin{tabular}{|l|l|l|}
\hline
Model    & Convergence Time & Word Similarity \\ \hline
SGD & {11}              & 0.65            \\ \hline
SGD w/shuffling & {7}              & 0.66            \\ \hline
Adam & {3}              & 0.66            \\ \hline
Adam w/ shuffling & \textbf{2}              & 0.66      \\ \hline
\end{tabular}
\label{table:results}
\caption{Results on the text8 dataset with different optimizers and input shuffling}

\end{table}
\iffalse
\begin{figure}[h]
    \centering
            \includegraphics[scale=0.3]{images/results_adam_shuffle} 
    \caption{Training time Adam with input Shuffling}
    \label{fig:results_adam_shuffle}
\end{figure}
\fi
